---
title: "Theodorou_ILC2_paper_figures"
author: "Jiangyan Yu (jiangyan.yu@uni-bonn.de)"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_folding: hide
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
Note:
1) data was downloaded from this paper https://www.cell.com/cell-metabolism/fulltext/S1550-4131(22)00394-1?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS1550413122003941%3Fshowall%3Dtrue#sectitle0030

2) qs file was downloaded from GEO: GSE183290
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE183288

3) this script is simply executed in the docker.
docker run --rm -p 8787:8787 -e USER=jyu -e PASSWORD=jyu -e ROOT=TRUE -v /home/agschlitzer/sciebo/Projects2023/Theodorou_ILC2_ownerJYu:/home/jyu/rstudio jiangyanyu/jyu_r4.1.2:1.0

# general steps

```{r global_options}
knitr::opts_chunk$set(warning=FALSE, messgae=FALSE, fig.path='Figs/', results = "hide")
## fig.width=4, fig.height=4
```

## libraries

```{r}
library(Seurat)
library(ComplexHeatmap)
library(data.table)
# remotes::install_version("DescTools",version = "0.99.44")
# library(DescTools)
library(magrittr)
library(dplyr)
library(stringr)
library(ggplot2)
# install.packages("qs")
# library(qs)
```

# working directory

```{r}
working_dir = "/home/jyu/rstudio/"
# working_dir = "/home/yu.j/sciebo/Projects2023/Theodorou_ILC2_ownerJYu/"
working_dir = "/Users/jiangyanyu/sciebo/Projects2023/Theodorou_ILC2_ownerJYu/"
```

# load ilc2 cells

```{r}
yang_vwat_ilc2 = readRDS(file = paste0(working_dir,"/manuscript/data/yang_vwat_ilc2.rds"))
DimPlot(yang_vwat_ilc2, label = TRUE)
```

# fig1A: umap

## rename

```{r}
Idents(yang_vwat_ilc2) = "seurat_clusters"
yang_vwat_ilc2 = RenameIdents(yang_vwat_ilc2,
                              "5" = "Progenitor",
                              "1" = "Progenitor",
                              "4" = "Progenitor",
                              "3" = "Intermediate",
                              "0" = "Mature",
                              "2" = "Mature")

yang_vwat_ilc2$paper_type = Idents(yang_vwat_ilc2)
```

## plot

```{r}
cell_color = c("Progenitor"="#619CFF",
               "Intermediate"="#00BA38",
               "Mature"="#F8766D")
pdf(file = paste0(working_dir,"/manuscript/fig1a_umap.pdf"),width = 4,height = 3)

p=DimPlot(yang_vwat_ilc2,label = TRUE,group.by = "paper_type",cols = cell_color)
p = p +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.title = element_blank())
plot(p)

dev.off()
p
```

# fig1b:markers

```{r}
tmp_genes = c("Il18r1","Il7r","Gata3")

pdf(file = paste0(working_dir,"/manuscript/fig1b_markers.pdf"),width = 6,height = 2)
p= FeaturePlot(yang_vwat_ilc2,features = tmp_genes,order = TRUE,ncol = 3,pt.size = 0.1) &
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.position = "none")
plot(p)
dev.off()
p

```

# fig1c:trajectory

see figure3_trajectory.rmd
```{r}

```

# figure 1d:density

```{r}
p = DimPlot(yang_vwat_ilc2, reduction = "umap", dims = c(1,2),label = FALSE,split.by = "intervention_group")+
  # scale_color_manual(values = defined_colors)+
  theme_classic()
  

p1 = ggplot(subset(p$data,intervention_group == "SC"),aes(UMAP_1,UMAP_2,color=ident))+
  geom_point(size=0.7)+
  xlim(-60,55)+
  ylim(-30,50)+
  scale_color_manual(values = cell_color)+
  geom_density_2d(color="grey",adjust=1)+
  theme_classic()+
  labs(title = "Control diet")+
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))

p2 = ggplot(subset(p$data,intervention_group == "SH"),aes(UMAP_1,UMAP_2,color=ident))+
  geom_point(size=0.7)+
  xlim(-60,55)+
  ylim(-30,50)+
  scale_color_manual(values = cell_color)+
  geom_density_2d(color="grey",adjust=1)+
  theme_classic()+
  labs(title = "High fat diet")+
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))

pdf(file = paste0(working_dir,"/manuscript/fig1c_density.pdf"),width = 6,height = 3)
p1+p2
dev.off()
p1+p2

```

# figs1a: umap split by condition

```{r}
p = DimPlot(yang_vwat_ilc2, reduction = "umap", dims = c(1,2),label = FALSE,split.by = "intervention_group")+
  scale_color_manual(values = cell_color)+
  theme_classic()
  

p1 = ggplot(subset(p$data,intervention_group == "SC"),aes(UMAP_1,UMAP_2,color=ident))+
  geom_point(size=0.7)+
  # xlim(-60,55)+
  # ylim(-30,50)+
  scale_color_manual(values = cell_color)+
  # geom_density_2d(color="grey",adjust=1)+
  theme_classic()+
  labs(title = "Control diet")+
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.position = "none",
        plot.title = element_text(hjust = 0.5))

# p1 = Seurat::LabelClusters(p1,id="seurat_clusters",color = "black")

p2 = ggplot(subset(p$data,intervention_group == "SH"),aes(UMAP_1,UMAP_2,color=ident))+
  geom_point(size=0.7)+
  # xlim(-60,55)+
  # ylim(-30,50)+
  scale_color_manual(values = cell_color)+
  # geom_density_2d(color="grey",adjust=1)+
  theme_classic()+
  labs(title = "High fat diet")+
  labs(color='Cell type') +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        # legend.position = "none",
        plot.title = element_text(hjust = 0.5))

# p2 = Seurat::LabelClusters(p2,id="seurat_clusters",color="black")

pdf(file = paste0(working_dir,"/manuscript/figs1a_umap_by_condition.pdf"),width = 4,height = 2)

# ggpubr::ggarrange(plotlist = list(p1,p2),ncol = 2,widths = c(1.3,1.9))
p1+p2
dev.off()
p1+p2
```




#figs1b:DEG

```{r}
# yang_vwat_ilc2_deg = FindAllMarkers(yang_vwat_ilc2)
# write.csv(yang_ilc2tnk_vwat_deg,file = paste0(working_dir,"/manuscript/data/yang_ilc2tnk_vwat.deg.csv"))

yang_vwat_ilc2_deg = read.csv(file = paste0(working_dir,"/manuscript/data/yang_vwat_ilc2_deg.csv"))

top10 = yang_vwat_ilc2_deg %>%
  subset(avg_log2FC >1) %>%
  group_by(cluster) %>%
  top_n(n=10,wt=avg_log2FC)

## reorder clusters
cell_order = c(4,5,1,3,2,0)
yang_vwat_ilc2$seurat_clusters = factor(yang_vwat_ilc2$seurat_clusters,levels = cell_order)
top10$cluster = factor(top10$cluster,levels = cell_order)
top10 = top10[order(top10$cluster),]

pdf(file = paste0(working_dir,"/manuscript/figs1b_deg.pdf"),height = 6,width = 4)
p=DoHeatmap(yang_vwat_ilc2,features = top10$gene,label = FALSE,group.colors  = cell_color)
p = p +scale_color_manual(values = cell_color)
plot(p)
dev.off()
p
# rm(top10)
```

# figs1c: bar plot

```{r}
ns <- table(sample = yang_vwat_ilc2$intervention_group, cell_type = yang_vwat_ilc2$paper_type)
fq <- prop.table(ns, 1) * 100
df <- as.data.frame(fq)
df = subset(df,sample == "SC" | sample == "SH")

df$sample = ifelse(df$sample == "SC","Control diet","High fat diet")
df$sample = factor(df$sample,levels = c("High fat diet","Control diet"))
df$cell_type = factor(df$cell_type,levels = c("Mature","Intermediate","Progenitor"))
# pdf(file = paste0(working_dir,"/manuscript/figs1c_bar.pdf"),width = 5, height = 2,colormodel="cmyk")
pdf(file = paste0(working_dir,"/manuscript/figs1c_bar.pdf"),width = 5, height = 2)


p = ggplot(df,aes(x=Freq,y=sample,fill=cell_type,label=cell_type))+
  geom_bar(stat="identity",colour="black")+
  scale_fill_manual(values=cell_color)+
  theme_classic()+
  labs(fill='Cell type') +
  # geom_text(position = position_stack(vjust = 0.5))+
  theme(axis.title=element_blank())

plot(p)
dev.off()
p
# rm(ns,fq,df,cluster_color,ms_folder)
```


# session info

```{r}
sessionInfo()
```

