---
title: "WAT_ILC2_trajectory_analysis"
author: "Jiangyan Yu (jiangyan.yu@uni-bonn.de)"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_folding: hide
    number_sections: yes
    theme: united
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
Note:
1) data was downloaded from this paper https://www.cell.com/cell-metabolism/fulltext/S1550-4131(22)00394-1?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS1550413122003941%3Fshowall%3Dtrue#sectitle0030

2) qs file was downloaded from GEO: GSE183290
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE183288

3) this script is simply executed in the docker.
docker run --rm -p 8787:8787 -e USER=jyu -e PASSWORD=jyu -e ROOT=TRUE -v /home/agschlitzer/sciebo/Projects2023/Theodorou_ILC2_ownerJYu:/home/jyu/rstudio jiangyanyu/jyu_r4.1.2:1.0

# general steps

```{r global_options}
knitr::opts_chunk$set(warning=FALSE, messgae=FALSE, fig.path='Figs/', results = "hide")
## fig.width=4, fig.height=4
```

## libraries

```{r}
library(Seurat)
library(ComplexHeatmap)
library(data.table)
remotes::install_version("DescTools",version = "0.99.44")
library(DescTools)
library(magrittr)
library(dplyr)
library(stringr)
library(ggplot2)
install.packages("qs")
library(qs)
```

# working directory

```{r}
working_dir = "/home/jyu/rstudio/"
```

# load yang published dataset

```{r}

yang_data = qread(file = paste0(working_dir,"/GSE183288_Single_cell_atlas.qs"))

yang_data

DimPlot(yang_data,raster = TRUE,label = TRUE)
```

# extract ILC2/T/NK

```{r}
yang_ilc2tnk = subset(yang_data,idents = c("NK","T","ILC"))
rm(yang_data)
gc()

DimPlot(yang_ilc2tnk,raster = TRUE)

# saveRDS(yang_ilc2tnk,file = paste0(working_dir,"/manuscript/data/yang_ilc2tnk_20230616.rds"))
```

## re-do seurat analysis 

```{r}
## select only vWAT cells

yang_ilc2tnk_vwat = subset(yang_ilc2tnk,subset=tissue=="vWAT")

rm(yang_ilc2tnk)
gc()

## redo umap and clustering analysis

yang_ilc2tnk_vwat = NormalizeData(yang_ilc2tnk_vwat)
yang_ilc2tnk_vwat = FindVariableFeatures(yang_ilc2tnk_vwat,selection.method = "vst",nfeatures = 2000)
yang_ilc2tnk_vwat = ScaleData(yang_ilc2tnk_vwat)
yang_ilc2tnk_vwat = RunPCA(yang_ilc2tnk_vwat,features = VariableFeatures(yang_ilc2tnk_vwat))
yang_ilc2tnk_vwat = FindNeighbors(yang_ilc2tnk_vwat,dims = 1:30)
yang_ilc2tnk_vwat = FindClusters(yang_ilc2tnk_vwat,resolution = 0.5)

yang_ilc2tnk_vwat = RunUMAP(yang_ilc2tnk_vwat,n.components = 3,n.neighbors = 30,dims = 1:50,min.dist = 0.1)
# yang_ilc2tnk_vwat = readRDS(file = paste0(working_dir,"/yang_ilc2tnk_vwat_20230712.rds"))

# DimPlot(yang_ilc2tnk_vwat,group.by = "seurat_clusters",label = TRUE)
DimPlot(yang_ilc2tnk_vwat,group.by = "seurat_clusters",label = TRUE,dims = c(1,3))
# DimPlot(yang_ilc2tnk_vwat,group.by = "seurat_clusters",label = TRUE,dims = c(2,3))

# DimPlot(yang_ilc2tnk_vwat,group.by = "cell_state_label",label = TRUE)
DimPlot(yang_ilc2tnk_vwat,group.by = "cell_state_label",label = TRUE,dims = c(1,3))
# DimPlot(yang_ilc2tnk_vwat,group.by = "cell_state_label",label = TRUE,dims = c(2,3))
```

## DEG heatmap

```{r,fig.width=8,fig.height=20}
Idents(yang_ilc2tnk_vwat) = "seurat_clusters"
# yang_ilc2tnk_vwat_deg = FindAllMarkers(yang_ilc2tnk_vwat)

# write.csv(yang_ilc2tnk_vwat_deg,file = paste0(working_dir,"/yang_ilc2tnk_vwat_deg.csv"))

yang_ilc2tnk_vwat_deg = read.csv(file = paste0(working_dir,"/yang_ilc2tnk_vwat_deg.csv"))

# saveRDS(yang_ilc2tnk_vwat,file = paste0(working_dir,"/yang_ilc2tnk_vwat_20230712.rds"))

top10 = yang_ilc2tnk_vwat_deg %>%
  group_by(cluster) %>%
  top_n(n=10,wt=avg_log2FC)

tmp_seu = yang_ilc2tnk_vwat@meta.data
tmp_seu$cell = rownames(tmp_seu)
tmp_seu = tmp_seu %>% group_by(seurat_clusters) %>% sample_n(40) 
tmp_seu = yang_ilc2tnk_vwat[,tmp_seu$cell]


DoHeatmap(tmp_seu,features = top10$gene,raster = TRUE)

rm(top10,tmp_seu)
```


## rename clusters

```{r}
Idents(yang_ilc2tnk_vwat) = "seurat_clusters"
yang_ilc2tnk_vwat = RenameIdents(yang_ilc2tnk_vwat,
                                 c("0" = "0:NKcell",
                                   "1" = "1:Tcell",
                                   "2" = "2:ILC2",
                                   "3" = "3:NK-T",
                                   "4" = "4:ILCP",
                                   "5" = "5:NKcell",
                                   "6" = "6:Tregcell",
                                   "7" = "7:Tcell",
                                   "8" = "8:Tcell",
                                   "9" = "9:inte-ILCP",
                                   "10" = "10:nd",
                                   "11" = "11:nd",
                                   "12" = "12:nd",
                                   "13" = "13:nd"))
yang_ilc2tnk_vwat$celltype1 = Idents(yang_ilc2tnk_vwat)
Idents(yang_ilc2tnk_vwat) = "celltype1"
```


## T, NK markers

```{r}
FeaturePlot(yang_ilc2tnk_vwat,features = "Cd3e",dims = c(1,3),order = TRUE)
FeaturePlot(yang_ilc2tnk_vwat,features = "Cd4",dims = c(1,3),order = TRUE)
FeaturePlot(yang_ilc2tnk_vwat,features = "Cd8a",dims = c(1,3),order = TRUE)
FeaturePlot(yang_ilc2tnk_vwat,features = "Klrb1c",dims = c(1,3),order = TRUE)
FeaturePlot(yang_ilc2tnk_vwat,features = "Foxp3",dims = c(1,3),order = TRUE)

DotPlot(yang_ilc2tnk_vwat,features = c("Cd3e","Cd4","Cd8a","Klrb1c","Foxp3"))+
  coord_flip()+
  theme(axis.text.x = element_text(
    angle = 90,hjust = 1,vjust = 0.5
  ))
# FeaturePlot(yang_ilc2tnk_vwat,features = "Roc",dims = c(1,3),order = TRUE)
```

## ILC markers

```{r}
tmp_genes = c("Il7r","Il18r1","Tbx21","Ikzf3","Eomes","Kir2dl3","Gata3","Maf","Ptgdr2","Hpgds","Rorc","Il23r","Il1r1","Kit")

DotPlot(yang_ilc2tnk_vwat,features = tmp_genes)+
  coord_flip()+
  theme(axis.text.x = element_text(
    angle = 90,hjust = 1,vjust = 0.5
  ))
```

# extract ILC2 cells

## re-do seurat analysis
```{r}
Idents(yang_ilc2tnk_vwat) = "seurat_clusters"

yang_vwat_ilc2 = subset(yang_ilc2tnk_vwat,idents=c(4,9,2))
yang_vwat_ilc2$tnkilc2_cluster = yang_vwat_ilc2$seurat_clusters

# DimPlot(yang_vwat_ilc2,label = TRUE,split.by = "intervention_group",ncol=2)
# 
yang_vwat_ilc2 = NormalizeData(yang_vwat_ilc2)
yang_vwat_ilc2 = FindVariableFeatures(yang_vwat_ilc2,selection.method = "vst",nfeatures = 2000)
yang_vwat_ilc2 = ScaleData(yang_vwat_ilc2)
yang_vwat_ilc2 = RunPCA(yang_vwat_ilc2,features = VariableFeatures(yang_vwat_ilc2))
yang_vwat_ilc2 = FindNeighbors(yang_vwat_ilc2,dims = 1:30)
yang_vwat_ilc2 = FindClusters(yang_vwat_ilc2,resolution = 0.5)

yang_vwat_ilc2 = RunUMAP(yang_vwat_ilc2,dims = 1:50,
                         n.components = 3,
                         min.dist = 1,
                         a = 0.01,
                         b = 1)

DimPlot(yang_vwat_ilc2,group.by = "tnkilc2_cluster",label = TRUE)

DimPlot(yang_vwat_ilc2,group.by = "seurat_clusters",label = TRUE)
# DimPlot(yang_vwat_ilc2,group.by = "seurat_clusters",label = TRUE,dims = c(1,3))
# DimPlot(yang_vwat_ilc2,group.by = "seurat_clusters",label = TRUE,dims = c(2,3))
```

## umap split by condition

```{r}
DimPlot(yang_vwat_ilc2,group.by = "seurat_clusters",label = TRUE,split.by = "intervention_group",ncol=2)

```

## DEG
```{r}
yang_vwat_ilc2_deg = FindAllMarkers(yang_vwat_ilc2)
write.csv(yang_vwat_ilc2_deg,file = paste0(working_dir,"/manuscript/data/yang_vwat_ilc2_deg.csv"))
# yang_ilc2tnk_vwat_deg = read.csv(file = paste0(working_dir,"/manuscript/data/yang_ilc2tnk_vwat.deg.csv"))
top10 = yang_vwat_ilc2_deg %>%
  subset(avg_log2FC >1) %>%
  group_by(cluster) %>%
  top_n(n=30,wt=avg_log2FC)


DoHeatmap(yang_vwat_ilc2,features = top10$gene)
```

    
# session infor
```{r}
saveRDS(yang_vwat_ilc2,file = paste0(working_dir,"/manuscript/data/yang_vwat_ilc2.rds"))
sessionInfo()
```

